Synthesizing Symmetric Lenses

### 2019-02-19 papers/SymmetricLenses/main.tex
update

To characterize their expressive power formally, we show that simple
symmetric lenses are exactly the symmetric lenses that satisfy an intuitive
property called \emph{forgetfulness}.  To check that they are expressive
enough in practice, we add simple symmetric lenses to the Boomerang
language~\cite{Boomerang} and apply them to a range of real-world
applications, showing in the process that simple symmetric lenses integrate
smoothly with other advanced lens features provided by Boomerang, including
quotient lenses~\cite{quotientlenses,maina+:quotient-synthesis} and matching
lenses~\cite{matchinglenses}.

@@

We now give typing rules and semantics for (some of) these combinators. We do not
include the semantics for combinators equivalent to those presented in prior
work~\cite{Boomerang}. Furthermore, we present
just \CreateR and \PutR in cases where \CreateL and \PutL are symmetric;
full details can be found in an appendix\ifappendices \S\ref{sec:appendixlenses}  \else{} (in the supplemental materials)\fi. For
simplicity of presentation, we use unambiguous regular expressions
everywhere, even though the identity and disconnect lenses can actually be defined over ambiguous
regular expressions; we believe relaxing
this restriction is unproblematic.

@@

The \MergeR lens is interesting because it merges data where one data can be in
two formats, and one data has only one format. In previous
work~\cite{Boomerang}, this was combined into \OrLens{}, where
\OrLens{} could have ambiguous types, but we find it more clear to have explicit
merge operators: it is easier to see what lens the synthesis algorithm is
creating.

>>>>

To characterize their expressive power formally, we show that simple
symmetric lenses are exactly the symmetric lenses that satisfy an intuitive
property called \emph{forgetfulness}.  To check that they are expressive
enough in practice, we add simple symmetric lenses to the Boomerang
language~\cite{boomerang} and apply them to a range of real-world
applications, showing in the process that simple symmetric lenses integrate
smoothly with other advanced lens features provided by Boomerang, including
quotient lenses~\cite{quotientlenses,maina+:quotient-synthesis} and matching
lenses~\cite{matchinglenses}.

@@

We now give typing rules and semantics for (some of) these combinators. We do not
include the semantics for combinators equivalent to those presented in prior
work~\cite{boomerang}. Furthermore, we present
just \CreateR and \PutR in cases where \CreateL and \PutL are symmetric;
full details can be found in an appendix\ifappendices \S\ref{sec:appendixlenses}  \else{} (in the supplemental materials)\fi. For
simplicity of presentation, we use unambiguous regular expressions
everywhere, even though the identity and disconnect lenses can actually be defined over ambiguous
regular expressions; we believe relaxing
this restriction is unproblematic.

@@

The \MergeR lens is interesting because it merges data where one data can be in
two formats, and one data has only one format. In previous
work~\cite{boomerang}, this was combined into \OrLens{}, where
\OrLens{} could have ambiguous types, but we find it more clear to have explicit
merge operators: it is easier to see what lens the synthesis algorithm is
creating.
